---
title: "Analysing Literacy Rate Changes in Indian States and Union Territories (2001–2011 census)"
author: "Japneet Kaur Oberoi" 
date: 10/04/2024
listing:
  contents: posts
  sort: "date desc"
  type: default
  categories: true
  sort-ui: false
  filter-ui: false
page-layout: full
title-block-banner: true
---


```{r, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
```

## Problem Description 

Any nation's literacy rate is a key sign of its socio-economic development. The actual literacy rate in India is calculated by dividing the total number of literate people aged 7 or older by the total population aged 7 or older, then multiplying the result by 100. Literacy data is gathered every ten years through the census. This regulates the percentage of the population proficient in reading and writing. In this blog we are looking at the variations in literacy rates between the 2001 and 2011 census periods across Indian states and union territories for the analysis. The objective is to comprehend local advancements and pinpoint areas that would require additional educational policies.


## Data Description

**Data Source**

The data set was downloaded from the [Kaggle Website](https://www.kaggle.com/) called [Govt Of India Literacy Rate](https://www.kaggle.com/datasets/doncorleone92/govt-of-india-literacy-rate?resource=download). Firstly we logged into the Kaggle website using a gmail account then we looked for literacy rates of India released by the government. The website provides a csv file which is easy to download and then it was loaded in Rstudio using the `read_csv` function.


```{r}
data <- read_csv("data/GOI.csv")
```

**Data Dictionary**

The data set has 8 variables which include information about the percentage of literacy rates of India and the states and Union territories in India. It also calculates the rates for Urban and rural populations for the years 2001 and 2011.
The variables are:

1) Category: Indicates whether the entry is a State or a Union Territory.
2) Country/ States/ Union Territories Name: The names of the state or union territory.
3) Literacy Rate (Persons) - Total - 2001: Overall percentage of literacy rate in 2001.
4) Literacy Rate (Persons) - Total - 2011: Overall percentage of literacy rate in 2011.
5) Literacy Rate (Persons) - Rural - 2001: Percentage of Rural literacy rate in 2001.
6) Literacy Rate (Persons) - Rural - 2011: Percentage of Rural literacy rate in 2011.
7) Literacy Rate (Persons) - Urban - 2001: Percentage of Urban literacy rate in 2001.
8) Literacy Rate (Persons) - Urban - 2011: Percentage of Urban literacy rate in 2011.


India has 28 States and 8 Union Territories. States enjoy a higher degree of autonomy, managing local governance, education, health, and law enforcement, while union territories are typically governed by the central government. Some union territories, like Delhi and Puducherry, have partial statehood, allowing them to have their own legislatures with limited authority. 

**Details of work**




## Analysis




